---
title: "TERN - morphology data for SLGA II - Get Data"
output: html_document
author: "Lauren O'Brien"
date: `r Sys.Date()`
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'hold', fig.height = 6,
                      fig.align = 'center', warning = FALSE, message = FALSE,
                      eval = TRUE)
```

```{r pkgs, results = 'hide'}
options(stringsAsFactors = FALSE, scipen = 999)
library(sf)            # vector data IO and processing
library(DBI)           # database interface
library(ROracle)       # Oracle database drivers
library(getPass)       # password management
library(units)         # for working with unit-attributed data
library(tidyverse)     # general data manipulation and display
library(RSQLite)       # local db write

Sys.setenv(
  'TNS_ADMIN'   = file.path('C:', 'Users', Sys.getenv('USERNAME'), 'Oracle'),
  'ORACLE_HOME' = 'C:/Program Files/Oracle Instant Client'
  )
SALI <- dbConnect(dbDriver('Oracle'),
                  username = 'obrienle',
                  password = getPass(),
                  dbname   = 'sispda',
                  host     = 'ex02client01',
                  port     = 1521)

if(!dir.exists(file.path(getwd(), 'SALI_DATA'))) {
  dir.create(file.path(getwd(), 'SALI_DATA'))
}

```

## Data snapshot

Take a snapshot of SALI sites module to work with offline. Filter on data that's allowed out - from published projects where QC has been completed.

The data-holding tables in the sites module are structured as follows:

1 SIT_SITES
    2 SIT_SITE_LAND_USES
    3 SIT_SITE_MAN_PRACS
    4 SIT_ELEM_GEOMORPHS 
    5 SIT_PATT_GEOMORPHS 
    6 SIT_NOTES 
    7 SIT_OBSERVATIONS
        8  SIT_LOCATIONS
        9  SIT_SOIL_CLASSIFICATIONS
        10 SIT_DISTURBANCES 
        11 SIT_SURFACE_CONDITIONS
        12 SIT_MICRORELIEFS
        13 SIT_EROSIONS
        14 SIT_SURF_COARSE_FRAGS
        15 SIT_ROCK_OUTCROPS
        16 SIT_SUB_MINERAL_COMPS
        17 SIT_VEG_COMMUNITIES
            18 SIT_VEG_STRATA
            19 SIT_VEG_STRATA_SPECIES
        20 SIT_HORIZONS
            21 SIT_HORIZON_SUFFICES
            22 SIT_HORIZON_COLOURS
            23 SIT_MOTTLES
            24 SIT_COARSE_FRAGS
            25 SIT_STRUCTURES
            26 SIT_SEGREGATIONS
            27 SIT_FABRICS
            28 SIT_CUTANS
            29 SIT_CRACKS
            30 SIT_MACROPORES
            31 SIT_STRENGTHS
            32 SIT_PANS
            33 SIT_ROOTS
            34 SIT_FIELD_TESTS
        35 SIT_SAMPLES
            36 SIT_LAB_RESULTS

Some additional tables hold decodes (meanings) for the coded data and are extracted separately below.

```{r getdat}
today <- '20191026' # edit for fresh snapshot

sit_datafile <- file.path(getwd(), 'SALI_DATA', 
                          paste0('SALI_SIT-data_', today, '.rds'))

# restricts to projects that are allowed out
temp1 <- c("with temp1 as (
             select project_code from reg_projects 
             where site_qc_status = 'P'
             and avail_status = 'E') ")

if(file.exists(sit_datafile)) {
  sit_data <- readRDS(sit_datafile)
} else {
  sit_data <- 
    list(
      # 1 - Sites
      "SIT" = 
        dbGetQuery(SALI,
         paste0(temp1, 
                "select * from temp1
                 inner join sit_sites using (project_code)
                 order by project_code, site_id")),
      # 2 - Land Uses
      # Contemporary to date(s) of observation(s) - these are not updated
      "SLU" = 
        dbGetQuery(SALI,
         paste0(temp1, 
                "select * from temp1
                 inner join sit_site_land_uses using (project_code)
                 order by project_code, site_id, land_use_no")),
      # 3 site management practices
      "SMP" = 
        dbGetQuery(SALI,
         paste0(temp1, 
                "select * from temp1
                 inner join sit_site_man_pracs using (project_code)
                 order by project_code, site_id, man_prac_no")),
      # 4 Geomorphology at element scale (radius 20m around site)
      "SGE" = 
        dbGetQuery(SALI,
         paste0(temp1, 
                "select * from temp1
                 inner join sit_elem_geomorphs using (project_code)
                 order by project_code, site_id, elem_geo_no")),
      # 5 Geomorphology at pattern scale (radius 300m around site)
      "SGP" = 
        dbGetQuery(SALI,
         paste0(temp1, 
                "select * from temp1
                 inner join sit_patt_geomorphs using (project_code)
                 order by project_code, site_id, patt_geo_no")),   
      # 6 Notes
      # NB Site, Obs and Hor notes all live in the one table but can be
      # keyed out easily by note_type
      # CFD notes are excluded as they contain names/phone numbers/addresses
      "NOT" = 
        dbGetQuery(SALI,
         paste0(temp1, 
                "select * from temp1
                 inner join sit_notes using (project_code)
                 where note_type != 'CFD'
                 order by project_code, site_id")),
      # 7 - Observations
      "OBS" = 
        dbGetQuery(SALI,
          paste0(temp1, 
            "select * from temp1
             inner join sit_observations using (project_code)
             order by project_code, site_id, obs_no")),
      # 8 Locations
      # grabbing all but SHAPE col below, ROracle can't handle SDO_GEOM :'(
      # also only grabbing coordinates in GDA94 (EPSG:4283). Some of these are
      # converted in the system, having originally been recorded in AGD84 or 
      # AGD66 datums, often in projected systems (commonly UTM zone 55S or 56S)
      "OLC" = 
        dbGetQuery(SALI,
          paste0(temp1, 
            "select project_code, site_id, obs_no, datum, derived_location,
             zone, easting, northing, longitude, latitude, loc_accuracy, 
             created_by, creation_date, last_updated_by, last_update_date
             from temp1
             inner join sit_locations using (project_code)
             where datum = 3
             and longitude > 1 and longitude is not null  /* >:-( */
             order by project_code, site_id, obs_no")),
      # 9 Soil Classifications
      "OCL" = 
        dbGetQuery(SALI,
          paste0(temp1, 
            "select * from temp1
             inner join sit_soil_classifications using (project_code)
             order by project_code, site_id, obs_no, soil_class_no")), 
      # 10 Disturbances
      "ODS" = 
        dbGetQuery(SALI,
          paste0(temp1, 
                 "select * from temp1
                  inner join sit_disturbances using (project_code)
                  order by project_code, site_id, obs_no, disturb_no")),
      # 11 surface conditions
      "OSC" = 
        dbGetQuery(SALI,
          paste0(temp1,
                 "select * from temp1
                  inner join sit_surf_conditions using (project_code)
                  order by project_code, site_id, obs_no, surf_cond_no")),
      # 12 Microrelief
      "OMC" = 
        dbGetQuery(SALI,
          paste0(temp1,
                 "select * from temp1
                  inner join sit_microreliefs using (project_code)
                  order by project_code, site_id, obs_no, micro_no")),
      # 13 Erosion
      "OER" = 
        dbGetQuery(SALI,
          paste0(temp1, 
                 "select * from temp1
                  inner join sit_erosions using (project_code)
                  order by project_code, site_id, obs_no, ero_no")),
      # 14 Surface Coarse Frags
      "OCF" =
        dbGetQuery(SALI,
          paste0(temp1,
                 "select * from temp1
                  inner join sit_surf_coarse_frags using (project_code)
                  order by project_code, site_id, obs_no, surf_frag_no")),
      # 15 Outcrop
      "OTC" = 
        dbGetQuery(SALI,
          paste0(temp1, 
                 "select * from temp1
                  inner join sit_rock_outcrops using (project_code)
                  order by project_code, site_id, obs_no, outcrop_no")),
      # 16 substrate mineral composition
      "OSM" = 
        dbGetQuery(SALI,
          paste0(temp1, 
                 "select * from temp1
                  inner join sit_sub_mineral_comps using (project_code)
                  order by project_code, site_id, obs_no, sub_min_no")),
      # 17 Vegetation communities
      "VCM" = 
        dbGetQuery(SALI,
          paste0(temp1, 
                 "select * from temp1
                  inner join sit_veg_communities using (project_code)
                  order by project_code, site_id, obs_no, veg_comm_no")),
      # 18 Veg strata
      "VST" =
        dbGetQuery(SALI,
          paste0(temp1,
                 "select * from temp1
                  inner join sit_veg_strata using (project_code)
                  order by project_code, site_id, obs_no, veg_comm_no")),
      # 19 veg spp
      "VSP" = 
        dbGetQuery(SALI,
            paste0(temp1, 
                   "select * from temp1
                    inner join sit_veg_strata_species using (project_code)
                    order by project_code, site_id, obs_no, veg_comm_no, 
                      veg_strata_spec_no")), 
      # 20 soil horizons
      "HOR" = 
        dbGetQuery(SALI,
          paste0(temp1, 
            "select hor.*, (texture_code || texture_qualifier) as texture 
             from temp1
             inner join sit_horizons hor 
               on temp1.project_code = hor.project_code
             order by hor.project_code, site_id, obs_no, horizon_no")),
      # 21 horizon suffices
      "HSX" = 
        dbGetQuery(SALI,
          paste0(temp1,
                 "select * from temp1
                  inner join sit_horizon_suffices using (project_code)
                  order by project_code, site_id, obs_no, horizon_no")),
      # 22 Horizon colours
      "HCO" = 
        dbGetQuery(SALI,
          paste0(temp1, 
                 "select * from temp1
                  inner join sit_horizon_colours using (project_code)
                  order by project_code, site_id, obs_no, horizon_no, 
                    hor_col_no")),
      # 23 mottles
      "HMT" = 
        dbGetQuery(SALI,
          paste0(temp1, 
                 "select * from temp1
                  inner join sit_mottles using (project_code)
                  order by project_code, site_id, obs_no, horizon_no, 
                    mott_no")),
      # 24 coarse frags
      "HCF" = 
        dbGetQuery(SALI,
         paste0(temp1, 
                "select * from temp1
                inner join sit_coarse_frags using (project_code)
                order by project_code, site_id, obs_no, horizon_no,
                  coarse_frag_no")),
      # 25 Stuctures
      "HST" = 
        dbGetQuery(SALI,
            paste0(temp1, 
                   "select * from temp1
                    inner join sit_structures using (project_code)
                    order by project_code, site_id, obs_no, horizon_no, 
                      struct_no")),    
      # 26 Segregations
      "HSG" =
        dbGetQuery(SALI,
            paste0(temp1, 
                   "select * from temp1
                    inner join sit_segregations using (project_code)
                    order by project_code, site_id, obs_no, horizon_no, 
                      seg_no")), 
      # 27 Fabrics
      "HFB" = 
        dbGetQuery(SALI,
          paste0(temp1, 
                 "select * from temp1
                  inner join sit_fabrics using (project_code)
                  order by project_code, site_id, obs_no, horizon_no, 
                    fab_no")),  
      # 28 Cutans
      "HCU" = 
        dbGetQuery(SALI,
          paste0(temp1, 
                 "select * from temp1
                  inner join sit_cutans using (project_code)
                  order by project_code, site_id, obs_no, horizon_no, 
                    cutan_no")),      
      # 29 Cracks
      "HCR" = 
        dbGetQuery(SALI,
          paste0(temp1, 
                 "select * from temp1
                  inner join sit_cracks using (project_code)
                  order by project_code, site_id, obs_no, horizon_no, 
                    crack_no")),
      # 30 Macropores
      "HMP" = 
        dbGetQuery(SALI,
            paste0(temp1, 
                   "select * from temp1
                    inner join sit_macropores using (project_code)
                    order by project_code, site_id, obs_no, horizon_no, 
                      macro_no")),  
      # 31 Strengths
      "HGT" =
        dbGetQuery(SALI,
            paste0(temp1, 
                   "select * from temp1
                    inner join sit_strengths using (project_code)
                    order by project_code, site_id, obs_no, horizon_no, 
                      stren_no")),
      # 32 Pans
      "HPA" =
        dbGetQuery(SALI,
            paste0(temp1, 
                   "select * from temp1
                    inner join sit_pans using (project_code)
                    order by project_code, site_id, obs_no, horizon_no, 
                      pan_no")),  
      # 33 Roots
      "HRT" =
        dbGetQuery(SALI,
            paste0(temp1, 
                   "select * from temp1
                    inner join sit_roots using (project_code)
                    order by project_code, site_id, obs_no, horizon_no, 
                      root_no")),   
      # 34 Field tests
      "FTS" = 
         dbGetQuery(SALI,
          paste0(temp1, 
                 "select * from temp1
                  inner join sit_field_tests using (project_code)
                  order by project_code, site_id, obs_no, horizon_no, 
                    test_no")),
      # 35 Samples
      # note - can join with keys including horizon but don't have to
      "SAM" = 
        dbGetQuery(SALI,
          paste0(temp1,
                 "select * from temp1
                  inner join sit_samples using (project_code)
                  order by project_code, site_id, obs_no, sample_no"))
      # 36 lab results 
      # see API at https://soil-chem.information.qld.gov.au/
      )
  saveRDS(sit_data, sit_datafile)
}

```

***

### Decode Tables

Decodes for short code lists are all in one large table called SALI_REF_CODES, grouped by field RV_DOMAIN

Longer code lists e.g. geology, landform elements are in their own tables

1 SALI_REF_CODES (used all over)
2 SIT_ELEM_TYPES 
3 SIT_PATT_TYPES 
4 SIT_GEOLOGIES  
5 SIT_LAND_USES 
6 SIT_LITHOLOGIES 
7 SIT_MASS_GEN_TYPES
8 SIT_ASCS
9 SIT_PPFS
10 SIT_GSGS
11 SIT_EROSION_TYPES
12 SIT_EROSION_DEGREES
13 SIT_VEG_SPECIES
14 SIT_STRUCT_FORMS
15 SIT_TEXTURES
16 SIT_COLOURS
17 SIT_COLOUR_CONVERT
18 SIT_FACIES
19 SIT_FIELD_TEST_TYPES
20 SIT_FIELD_TEST_TYPE_SPECS
21 SIT_NOTE_TYPES

```{r getdecodes}
decode_datafile <- file.path(getwd(), 'SALI_DATA', 
                             paste0('SALI_SIT-data_decodes_', today, '.rds'))

if(file.exists(decode_datafile)) {
  decodes <- readRDS(decode_datafile)
} else {
  decodes <- 
    list(
      # SALI ref codes holds all the short lists 
      "SRC" = 
        # not all of these domains are relevant but overkill to filter
        dbGetQuery(SALI,
                   "select * from sali_ref_codes
                    order by rv_domain, rv_low_value"
        ) ,
      "ELE" = dbGetQuery(SALI, "select * from sit_elem_types"),
      "PAT" = dbGetQuery(SALI, "select * from sit_patt_types"),
      "GEO" = dbGetQuery(SALI, "select * from sit_geologies"),
      "LUS" = dbGetQuery(SALI, "select * from sit_land_uses"),
      "LTH" = dbGetQuery(SALI, "select * from sit_lithologies"), 
      "MGT" = dbGetQuery(SALI, "select * from sit_mass_gen_types"), 
      "ASC" = dbGetQuery(SALI, "select * from sit_ascs"),
      "PPF" = dbGetQuery(SALI, "select * from sit_ppfs"),
      "GSG" = dbGetQuery(SALI, "select * from sit_gsgs"),
      "ERT" = dbGetQuery(SALI, "select * from sit_erosion_types"),      
      "ERD" = dbGetQuery(SALI, "select * from sit_erosion_degrees"),
      "VSS" = dbGetQuery(SALI, "select * from sit_veg_species"),
      "VST" = dbGetQuery(SALI, "select * from sit_struct_forms"),
      "TEX" = dbGetQuery(SALI, "select * from sit_textures"),
      "COL" = dbGetQuery(SALI, "select * from sit_colours"),
      "CLV" = dbGetQuery(SALI, "select * from sit_colour_convert"),
      "FAC" = dbGetQuery(SALI, "select * from sit_facies"),
      "FTT" = dbGetQuery(SALI, "select * from sit_field_test_types"),
      "FTS" = dbGetQuery(SALI, "select * from sit_field_test_type_specs"),
      "NTP" = dbGetQuery(SALI, "select * from sit_note_types"), 
    )
  saveRDS(decodes, decode_datafile)
}
      
```

### Make a multilayer geopackage

Because why not?

```{r biggp}
# make locations spatial
SIT_LOCATIONS <- st_as_sf(sit_data[['OLC']],
                          coords = c('LONGITUDE', 'LATITUDE'),
                          crs = 4283, remove = FALSE,
                          dim = 'XY')

# write locations to gpkg
st_write(SIT_LOCATIONS, 
         file.path(getwd(), 'SALI_DATA',
                   paste0('SALI_SITES_', format(Sys.Date(), '%Y%m%d'), '.gpkg')),
         layer = 'SIT_LOCATIONS',
         driver = 'GPKG', delete_dsn = TRUE)

# pop locations from main list and rename
sit_data[['OLC']] <- NULL

table_names <- c('SIT_SITES', 'SIT_SITE_LAND_USES', 'SIT_SITE_MAN_PRACS',
                 'SIT_ELEM_GEOMORPHS', 'SIT_PATT_GEOMORPHS', 'SIT_NOTES', 
                 'SIT_OBSERVATIONS', 'SIT_SOIL_CLASSIFICATIONS',
                 'SIT_DISTURBANCES', 'SIT_SURFACE_CONDITIONS', 'SIT_MICRORELIEFS', 
                 'SIT_EROSIONS',  'SIT_SURF_COARSE_FRAGS', 'SIT_ROCK_OUTCROPS', 
                 'SIT_SUB_MINERAL_COMPS', 'SIT_VEG_COMMUNITIES', 'SIT_VEG_STRATA', 
                 'SIT_VEG_STRATA_SPECIES', 'SIT_HORIZONS', 'SIT_HORIZON_SUFFICES', 
                 'SIT_HORIZON_COLOURS', 'SIT_MOTTLES', 'SIT_COARSE_FRAGS', 
                 'SIT_STRUCTURES', 'SIT_SEGREGATIONS', 'SIT_FABRICS', 'SIT_CUTANS',
                 'SIT_CRACKS', 'SIT_MACROPORES', 'SIT_STRENGTHS', 'SIT_PANS', 
                 'SIT_ROOTS', 'SIT_FIELD_TESTS', 'SIT_SAMPLES')
names(sit_data) <- table_names

# do the same for decode tables - note name change, should make the tables
# easier to work with. I always struggle to remember which tables 
# actually hold data vs decodes...
decode_names <- c('REF_CODES', 'REF_ELEM_TYPES', 'REF_PATT_TYPES', 
                  'REF_GEOLOGIES', 'REF_LAND_USES', 'REF_LITHOLOGIES', 
                  'REF_MASS_GEN_TYPES', 'REF_ASCS', 'REF_PPFS', 'REF_GSGS', 
                  'REF_EROSION_TYPES', 'REF_EROSION_DEGREES', 'REF_VEG_SPECIES', 
                  'REF_STRUCT_FORMS', 'REF_TEXTURES', 'REF_COLOURS', 
                  'REF_COLOUR_CONVERT', 'REF_FACIES', 'REF_FIELD_TEST_TYPES', 
                  'REF_FIELD_TEST_TYPE_SPECS')
names(decodes) <- decode_names

# add to gpkg
con <- dbConnect(SQLite(), 
                 dbname = file.path(getwd(),'SALI_DATA',
                                    paste0('SALI_SITES_',
                                           format(Sys.Date(), '%Y%m%d'),
                                           '.gpkg'))
                 )
purrr::map2(sit_data, table_names, function(lyr, nm) { dbWriteTable(con, nm, lyr) })
purrr::map2(decodes, decode_names, function(lyr, nm) { dbWriteTable(con, nm, lyr) })
dbDisconnect(con)
  
# about 300MB on disk so not v practical?
# Need to use Virtual Layer in QGIS (and whatever the Arc equiv is... make Query
# Layer??) to join tables
# easier to work with in R/Py, likely
```

```{r plotdemo, echo = FALSE, results = 'hold'}

sph_bab <- dplyr::filter(sit_data[['SIT_FIELD_TESTS']], PROJECT_CODE == 'BAB' &
                         TEST_TYPE == 'PH' & OBS_NO == 1 & 
                           HORIZON_NO  == 1 & TEST_NO == 1) %>% 
  left_join(., SIT_LOCATIONS, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  mutate(VALUE = as.numeric(VALUE),
         VALUE = factor(VALUE, levels = sort(unique(VALUE)), ordered = TRUE)) %>% 
  st_as_sf(crs = 4283)

phpal <- colorRampPalette(scales::brewer_pal('div', 'RdBu')(14))

ggplot() +
  geom_sf(data = sph_bab, aes(col = VALUE), pch = 16, show.legend = 'point') +
  scale_colour_manual(values = phpal(14)) +
  labs(col = 'pH') +
  ggtitle('Bundaberg 1:50,000 soil sites', subtitle = 'Surface Raupach pH') +
  theme_minimal()

```

***
