---
title: "Ranking Texture Codes"
author: "Lauren O'Brien"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hold', fig.height = 6,
                      fig.align = 'center', warning = FALSE, message = FALSE,
                      eval = FALSE)
```

Trying to introduce some ranking measures for soil texture codes. This allows us to do things like report the relative heaviest or lightest texture in a profile (or part of the solum).

```{r pkgs, results = 'hide'}
options(stringsAsFactors = FALSE, scipen = 999)
library(sf)            # vector data IO and processing
library(DBI)           # database interface
library(ROracle)       # Oracle database drivers
library(getPass)       # password management
library(units)         # for working with unit-attributed data
library(tidyverse)     # general data manipulation and display

Sys.setenv(
  'TNS_ADMIN'   = file.path('C:', 'Users', Sys.getenv('USERNAME'), 'Oracle'),
  'ORACLE_HOME' = 'C:/Program Files/Oracle Instant Client'
  )
SALI <- dbConnect(dbDriver('Oracle'),
                  username = 'obrienle',
                  password = getPass(),
                  dbname   = 'sispda',
                  host     = 'ex02client01',
                  port     = 1521)

if(!dir.exists(file.path(getwd(), 'SALI_DATA'))) {
  dir.create(file.path(getwd(), 'SALI_DATA'))
}

if(!dir.exists(file.path(getwd(), 'DERIVED_DATA'))) {
  dir.create(file.path(getwd(), 'DERIVED_DATA'))
}
```

There are a large number of distinct texture codes in use across SALI:

```{r gettc}
all_tcs <- dbGetQuery(SALI, 
  "select (texture_code || texture_qualifier) as texture, 
    texture_code, texture_qualifier, description, qual_desc,
    count(*) as n_codes
   from sit_horizons 
   left outer join sit_textures using (texture_code)
   left outer join (select rv_low_value as texture_qualifier, 
                      rv_meaning as qual_desc 
                    from sali_ref_codes 
                    where rv_domain = 'TEXTURE QUALIFIERS') src
    using(texture_qualifier)
   group by (texture_code || texture_qualifier), texture_code, texture_qualifier, 
     description, qual_desc
   order by texture_code, texture_qualifier")

```

Nearly 300 in fact, counting qualifiers.

The 'core' texture codes and their common modifiers (K for coarse, F for fine, Z for silty) are defined in the Yellow Book, and are nominally ordered by increasing overall fineness as such:

  * Sand: KS, S, FS,
  * Loamy Sand: LKS, LS, LFS
  * Clayey Sand: CKS, CS, CFS
  * Sandy Loam: KSL, SL, FSL, ZL
  * Loam: LSY, LFSY, L
  * Sandy Clay Loam: KSCL, SCL, FSCL, ZCL
  * Clay Loam: CLKS, CLS, CLFS, CLZ, CL
  * Light Clay: KSLC, SLC, FSLC, ZLC, LC
  * Light Medium Clay: KSLMC, SLMC, FSLMC, ZLMC, LMC
  * Medium Clay: KSMC, SMC, FSMC, ZMC, MC
  * Medium Heavy Clay: KSMHC, SMHC, FSMHC, ZMHC, MHC
  * Heavy Clay: KSHC, SHC, FSHC, ZHC, HC

Unmodified codes like LS are held to contain 'medium sand', the size range of which is not discussed in the Yellow Book. There are also some +/- modifiers that can be applied to non-clay textures (clay textures have them built in), and some A/P modifiers for mineral soils that are not quite peats but are highly organic. Peat codes exist but are kind of off to one side (IP, AP, HP) - but they'll have to be ranked in somehow, or just ignored.

SALI data is still messier than this though, because it contains historical data recorded pre-Yellow Book and pre-SALI, when people used different systems or (apparently) just made up codes as they went along. 

Thinking about a nested ranking with three levels:

  * Level 1 - 'clay', 'clay loam', 'loam', 'sand', 'gravel', 'peat'. Useful for things like suitability subclasses and layperson-oriented decodes.
  * Level 2 - splits classes up - light/medium/heavy for clays, light/heavy for the others
  * Level 3 - ranking within that on sand fraction fineness (off descriptions)
  * Level 4 - throw in modifiers
  
### Level 1

```{r plusminus}
main_tcs <- all_tcs %>% 
  filter(!is.na(TEXTURE_CODE)) %>% 
  select(TEXTURE_CODE, DESCRIPTION) %>% 
  distinct() %>% 
  arrange(TEXTURE_CODE) 

tc_ranks <- main_tcs %>% 
  mutate(GROUP = case_when(TEXTURE_CODE == 'AP'     ~ 'peat',
                           TEXTURE_CODE == 'CFS'    ~ 'sand',
                           TEXTURE_CODE == 'CKS'    ~ 'sand',
                           TEXTURE_CODE == 'CL'     ~ 'clay loam',
                           TEXTURE_CODE == 'CLFS'   ~ 'clay loam',
                           TEXTURE_CODE == 'CLFSZ'  ~ 'clay loam', 
                           TEXTURE_CODE == 'CLKS'   ~ 'clay loam',
                           TEXTURE_CODE == 'CLMS'   ~ 'clay loam',
                           TEXTURE_CODE == 'CLS'    ~ 'clay loam',
                           TEXTURE_CODE == 'CLZ'    ~ 'clay loam', 
                           TEXTURE_CODE == 'CP'     ~ 'peat',
                           TEXTURE_CODE == 'CS'     ~ 'sand',
                           TEXTURE_CODE == 'CSC'    ~ 'clay',
                           TEXTURE_CODE == 'FS'     ~ 'sand',
                           TEXTURE_CODE == 'FSC'    ~ 'clay',
                           TEXTURE_CODE == 'FSCL'   ~ 'loam',
                           TEXTURE_CODE == 'FSCLZ'  ~ 'loam', 
                           TEXTURE_CODE == 'FSHC'   ~ 'clay', 
                           TEXTURE_CODE == 'FSL'    ~ 'loam', 
                           TEXTURE_CODE == 'FSLC'   ~ 'clay', 
                           TEXTURE_CODE == 'FSLCZ'  ~ 'clay', 
                           TEXTURE_CODE == 'FSLMC'  ~ 'clay',
                           TEXTURE_CODE == 'FSLZ'   ~ 'loam',
                           TEXTURE_CODE == 'FSMC'   ~ 'clay',
                           TEXTURE_CODE == 'FSMHC'  ~ 'clay',
                           TEXTURE_CODE == 'GP'     ~ 'peat',
                           TEXTURE_CODE == 'GR'     ~ 'gravel',
                           TEXTURE_CODE == 'HC'     ~ 'clay',
                           TEXTURE_CODE == 'HP'     ~ 'peat',
                           TEXTURE_CODE == 'IP'     ~ 'peat',
                           TEXTURE_CODE == 'KS'     ~ 'sand',
                           TEXTURE_CODE == 'KSC'    ~ 'clay',
                           TEXTURE_CODE == 'KSCL'   ~ 'loam',
                           TEXTURE_CODE == 'KSHC'   ~ 'clay',
                           TEXTURE_CODE == 'KSL'    ~ 'loam',
                           TEXTURE_CODE == 'KSLC'   ~ 'clay',
                           TEXTURE_CODE == 'KSLMC'  ~ 'clay',
                           TEXTURE_CODE == 'KSMC'   ~ 'clay',
                           TEXTURE_CODE == 'KSMHC'  ~ 'clay',
                           TEXTURE_CODE == 'KSS'    ~ 'sand',
                           TEXTURE_CODE == 'L'      ~ 'loam',
                           TEXTURE_CODE == 'LC'     ~ 'clay',
                           TEXTURE_CODE == 'LCFS'   ~ 'clay',
                           TEXTURE_CODE == 'LCKS'   ~ 'clay',
                           TEXTURE_CODE == 'LCS'    ~ 'clay',
                           TEXTURE_CODE == 'LCZ'    ~ 'clay',
                           TEXTURE_CODE == 'LFS'    ~ 'sand',
                           TEXTURE_CODE == 'LFSY'   ~ 'loam',
                           TEXTURE_CODE == 'LFSYZ'  ~ 'loam',
                           TEXTURE_CODE == 'LKS'    ~ 'sand',
                           TEXTURE_CODE == 'LMC'    ~ 'clay',
                           TEXTURE_CODE == 'LMCFS'  ~ 'clay',
                           TEXTURE_CODE == 'LMCKS'  ~ 'clay',
                           TEXTURE_CODE == 'LMCS'   ~ 'clay',
                           TEXTURE_CODE == 'LMCZ'   ~ 'clay',
                           TEXTURE_CODE == 'LMS'    ~ 'sand',
                           TEXTURE_CODE == 'LP'     ~ 'peat',
                           TEXTURE_CODE == 'LS'     ~ 'sand',
                           TEXTURE_CODE == 'LSY'    ~ 'loam',
                           TEXTURE_CODE == 'MC'     ~ 'clay',
                           TEXTURE_CODE == 'MCFS'   ~ 'clay',
                           TEXTURE_CODE == 'MCS'    ~ 'clay',
                           TEXTURE_CODE == 'MCZ'    ~ 'clay',
                           TEXTURE_CODE == 'MHC'    ~ 'clay',
                           TEXTURE_CODE == 'MHCFS'  ~ 'clay',
                           TEXTURE_CODE == 'MHCS'   ~ 'clay',
                           TEXTURE_CODE == 'MS'     ~ 'sand',
                           TEXTURE_CODE == 'MSC'    ~ 'clay',
                           TEXTURE_CODE == 'MSCL'   ~ 'loam',
                           TEXTURE_CODE == 'S'      ~ 'sand',
                           TEXTURE_CODE == 'SC'     ~ 'clay',
                           TEXTURE_CODE == 'SCL'    ~ 'loam',
                           TEXTURE_CODE == 'SCLFS'  ~ 'loam',
                           TEXTURE_CODE == 'SHC'    ~ 'clay',
                           TEXTURE_CODE == 'SL'     ~ 'loam',
                           TEXTURE_CODE == 'SLC'    ~ 'clay',
                           TEXTURE_CODE == 'SLMC'   ~ 'clay',
                           TEXTURE_CODE == 'SLZ'    ~ 'loam',
                           TEXTURE_CODE == 'SMC'    ~ 'clay',
                           TEXTURE_CODE == 'SMHC'   ~ 'clay',
                           TEXTURE_CODE == 'SP'     ~ 'peat',
                           TEXTURE_CODE == 'SS'     ~ 'sand',
                           TEXTURE_CODE == 'ST'     ~ 'gravel',
                           TEXTURE_CODE == 'ZC'     ~ 'clay',
                           TEXTURE_CODE == 'ZCL'    ~ 'clay loam',
                           TEXTURE_CODE == 'ZHC'    ~ 'clay',
                           TEXTURE_CODE == 'ZL'     ~ 'loam',
                           TEXTURE_CODE == 'ZLC'    ~ 'clay',
                           TEXTURE_CODE == 'ZLCFS'  ~ 'clay',
                           TEXTURE_CODE == 'ZLMC'   ~ 'clay',
                           TEXTURE_CODE == 'ZLMCS'  ~ 'clay',
                           TEXTURE_CODE == 'ZMC'    ~ 'clay',
                           TEXTURE_CODE == 'ZMHC'   ~ 'clay',
                           TRUE ~ NA_character_)) %>% 
  mutate(GROUP = factor(GROUP, levels = c('gravel', 'sand', 'peat', 'loam',
                                          'clay loam', 'clay'), 
                        ordered = TRUE))
```

### Level 2

  * Gravel: no subdivision
  * Sand: clayey sand, loamy sand, sand
  * Peat: Minerally, not
  * Loam: sandy clay loam, loam, sandy loam, silty loam
  * Clay loam: sandy, silty, not
  * Clay: Light, light medium, medium, medium heavy, heavy 
  
```{r grp2spl}
# note putting unspecified clays like "coarse sandy clay CSC" in with LC group
# based on mapping across from Northcote

tc_ranks <- tc_ranks %>% 
  mutate(GROUP_2 = 
           case_when(GROUP == 'gravel' ~ 'gravel',
                     GROUP == 'sand' & 
                       TEXTURE_CODE %in% c('CFS', 'CKS', 'CS') ~ 'clayey sand',
                     GROUP == 'sand' & 
                       TEXTURE_CODE %in% c('LMS', 'LS', 'LKS', 'LFS') ~ 'loamy sand',
                     GROUP == 'sand' &
                       TEXTURE_CODE %in% c('S', 'SS', 'MS', 'FS', 'KS', 'KSS') ~ 'sand',
                     GROUP == 'peat' &
                       TEXTURE_CODE %in% c('CP', 'LP', 'SP') ~ 'mineral peat',
                     GROUP == 'peat' & 
                       TEXTURE_CODE %in% c('GP', 'IP', 'HP', 'AP') ~ 'peat',
                     GROUP == 'loam' &
                       TEXTURE_CODE %in% c('ZL') ~ 'silty loam',
                     GROUP == 'loam' &
                       TEXTURE_CODE %in% c('SL', 'SLZ', 'KSL', 'FSL', 'FSLZ') ~ 'sandy loam',
                     GROUP == 'loam' &
                       TEXTURE_CODE %in% c('FSCL', 'FSCLZ', 'KSCL', 'MSCL', 
                                           'SCL', 'SCLFS') ~ 'sandy clay loam',
                     GROUP == 'loam' &
                       TEXTURE_CODE %in% c('L', 'LFSY', 'LFSYZ', 'LSY') ~ 'loam',
                     GROUP == 'clay loam' &
                       TEXTURE_CODE %in% c('ZCL', 'CLZ') ~ 'clay loam, silty',
                     GROUP == 'clay loam' &
                       TEXTURE_CODE %in% c('CL') ~ 'clay loam',
                     GROUP == 'clay loam' &
                       TEXTURE_CODE %in% c('CLFS', 'CLFSZ', 'CLKS', 'CLMS', 'CLS') ~ 'clay loam, sandy',
                     GROUP == 'clay' &
                       TEXTURE_CODE %in% c('FSLC', 'FSLCZ', 'KSLC', 'LC', 
                                           'LCFS', 'LCKS', 'LCS', 'LCZ', 'SLC', 
                                           'ZLC', 'ZLCFS',
                                           # Northcote (?!?):
                                           'CSC', 'FSC', 'KSC', 'MSC', 'SC', 'ZC') ~ 'light clay',
                     GROUP == 'clay' & 
                       TEXTURE_CODE %in% c('FSLMC', 'KSLMC', 'LMC', 'LMCFS', 
                                           'LMCKS', 'LMCS', 'LMCZ', 'SLMC',
                                           'ZLMC', 'ZLMCS') ~ 'light medium clay',
                     
                     GROUP == 'clay' &
                       TEXTURE_CODE %in% c('FSMC', 'KSMC', 'MC', 'MCFS', 'MCS',
                                           'MCZ', 'SMC', 'ZMC') ~ 'medium clay',
                     GROUP == 'clay' &
                       TEXTURE_CODE %in% c('FSMHC', 'KSMHC', 'MHC', 'MHCFS', 
                                           'MHCS', 'SMHC', 'ZMHC') ~ 'medium heavy clay',
                     GROUP == 'clay' &
                       TEXTURE_CODE %in% c('FSHC', 'HC', 'KSHC', 'SHC', 'ZHC') ~ 'heavy clay',
                     TRUE ~ NA_character_
                     )) %>% 
  mutate(GROUP_2 = factor(GROUP_2,
                          levels = c('gravel', 
                                     'sand', 'loamy sand', 'clayey sand',
                                     'peat', 'mineral peat', 
                                     'sandy loam', 'loam', 'sandy clay loam', 'silty loam',
                                     'clay loam, sandy', 'clay loam, silty', 'clay loam',
                                     'light clay', 'light medium clay', 'medium clay', 
                                     'medium heavy clay', 'heavy clay'), 
                        ordered = TRUE)) %>% 
  arrange(GROUP, GROUP_2)

```

### Group 3

Advice from BH for 'normal vs silty' - put silty as heaviest within each group (congruent with limited data from PSAs)

  * Gravel: no subdivision
  * Sand: coarse, medium, fine
  * Loamy sand: coarse, medium, fine
  * Clayey sand: coarse, medium, fine
  * Peat: fibric, hemic, sapric  mapped to coarse, medium, fine. Granular is a wierd one - highly decomposed but agg'd to 1-2mm particle size, but would have to act differently to a sand at FC. Lumping with fibric for now.
  * Mineral Peat: sandy, loamy ,clayey mapped to coarse, medium, fine.
  * Sandy loam: coarse, medium, fine 
  * Loam: loam , medium sandy, fine sandy
  * Silty loam: no subdividision
  * Sandy Clay Loam: coarse, medium, fine
  * Clay loam: coarse, medium, fine
  * Clay loam, silty: no subsivision
  * clay loam: no subdivision
  * clays : coarse, medium, fine sandy, silty, no modifier
  
These start to look a lot like the DESCRIPTION field we already have.
  
```{r grp3spl}
tc_ranks <- tc_ranks %>% 
  mutate(GROUP_3 = 
           case_when(GROUP_2 == 'gravel' ~ 'gravel',
                     GROUP_2 == 'sand' &
                       TEXTURE_CODE %in% c('KS', 'KSS') ~ 'sand, coarse',
                     GROUP_2 == 'sand' &
                       TEXTURE_CODE %in% c('MS', 'S', 'SS') ~ 'sand, medium',
                     GROUP_2 == 'sand' &
                       TEXTURE_CODE %in% c('FS') ~ 'sand, fine',
                     GROUP_2 == 'loamy sand' &
                       TEXTURE_CODE %in% c('LKS') ~ 'loamy sand, coarse',
                     GROUP_2 == 'loamy sand' &
                       TEXTURE_CODE %in% c('LS' ,'LMS') ~ 'loamy sand, medium',
                     GROUP_2 == 'loamy sand' &
                       TEXTURE_CODE %in% c('LFS') ~ 'loamy sand, fine',
                     GROUP_2 == 'clayey sand' &
                       TEXTURE_CODE == 'CKS' ~ 'clayey sand, coarse',
                     GROUP_2 == 'clayey sand' &
                       TEXTURE_CODE == 'CS' ~ 'clayey sand, medium',
                     GROUP_2 == 'clayey sand' &
                       TEXTURE_CODE == 'CFS' ~ 'clayey sand, fine',
                     GROUP_2 == 'peat' &
                       TEXTURE_CODE %in% c('GP', 'IP') ~ 'peat, coarse',
                     GROUP_2 == 'peat' &
                       TEXTURE_CODE == 'HP' ~ 'peat, medium',
                     GROUP_2 == 'peat' &
                       TEXTURE_CODE == 'AP' ~ 'peat, fine',
                     GROUP_2 == 'mineral peat' &
                       # don't like these three...
                       TEXTURE_CODE == 'SP' ~ 'mineral peat, coarse',
                     GROUP_2 == 'mineral peat' &
                       TEXTURE_CODE == 'LP' ~ 'mineral peat, medium',
                     GROUP_2 == 'mineral peat' &
                       TEXTURE_CODE == 'CP' ~ 'mineral peat, fine',
                     GROUP_2 == 'sandy loam' &
                       TEXTURE_CODE %in% c('KSL') ~ 'sandy loam, coarse',
                     GROUP_2 == 'sandy loam' &
                       TEXTURE_CODE %in% c('SL')  ~ 'sandy loam, medium',
                     GROUP_2 == 'sandy loam' &
                       TEXTURE_CODE %in% c('FSL', 'FSLZ', 'SLZ') ~ 'sandy loam, fine',
                     GROUP_2 == 'loam' &
                       TEXTURE_CODE %in% c('L') ~ 'loam',
                     GROUP_2 == 'loam' &
                       TEXTURE_CODE %in% c('LFSY', 'LFSYZ') ~ 'loam, sandy, fine',
                     GROUP_2 == 'loam' &
                       TEXTURE_CODE %in% c('LSY') ~ 'loam, sandy',
                     GROUP_2 == 'silty loam' ~ 'silty loam',
                     GROUP_2 == 'sandy clay loam' &
                       TEXTURE_CODE %in% c('KSCL') ~ 'sandy clay loam, coarse',
                     GROUP_2 == 'sandy clay loam' &
                       TEXTURE_CODE %in% c('MSCL', 'SCL') ~ 'sandy clay loam, medium',
                     GROUP_2 == 'sandy clay loam' &
                       TEXTURE_CODE %in% c('FSCL', 'FSCLZ', 'SCLFS') ~ 'sandy clay loam, fine',
                     GROUP_2 == 'clay loam, sandy' &
                       TEXTURE_CODE %in% c('CLKS') ~ 'clay loam, sandy, coarse',
                     GROUP_2 == 'clay loam, sandy' &
                       TEXTURE_CODE %in% c('CLMS', 'CLS')  ~ 'clay loam, sandy, medium',
                     GROUP_2 == 'clay loam, sandy' &
                       TEXTURE_CODE %in% c('CLFS', 'CLFSZ') ~ 'clay loam, sandy, fine',
                     GROUP_2 == 'clay loam, silty' ~ 'clay loam, silty',
                     GROUP_2 == 'clay loam' ~ 'clay loam',
                     GROUP_2 == 'light clay' &
                       TEXTURE_CODE %in% c('CSC', 'KSC', 'KSLC', 'LCKS') ~ 'light clay, sandy, coarse',
                     GROUP_2 == 'light clay' &
                       TEXTURE_CODE %in% c('LCS', 'SC', 'SLC', 'MSC') ~ 'light clay, sandy, medium',
                     GROUP_2 == 'light clay' &
                       TEXTURE_CODE %in% c('FSC', 'FSLC', 'LCFS') ~ 'light clay, sandy, fine',
                     GROUP_2 == 'light clay' &
                       TEXTURE_CODE %in% c('LCZ', 'ZC', 'ZLC') ~ 'light clay, silty',
                     GROUP_2 == 'light clay' &
                       TEXTURE_CODE %in% c('LC', 'ZLCFS', 'FSLCZ') ~ 'light clay', # putting the 'I can't make up my mind' codes in here but that needs checking
                     GROUP_2 == 'light medium clay' &
                       TEXTURE_CODE %in% c('KSLMC', 'LMCKS') ~  'light medium clay, sandy, coarse',
                     GROUP_2 == 'light medium clay' &
                       TEXTURE_CODE %in% c('LMCS', 'SLMC') ~  'light medium clay, sandy, medium',
                     GROUP_2 == 'light medium clay' &
                       TEXTURE_CODE %in% c('FSLMC', 'LMCFS') ~  'light medium clay, sandy, fine',
                     GROUP_2 == 'light medium clay' &
                       TEXTURE_CODE %in% c('LMCZ', 'ZLMC') ~  'light medium clay, silty',
                     GROUP_2 == 'light medium clay' &
                       TEXTURE_CODE %in% c('LMC', 'ZLMCS') ~  'light medium clay',
                     GROUP_2 == 'medium clay' &
                       TEXTURE_CODE %in% c('KSMC') ~ 'medium clay, sandy, coarse',
                     GROUP_2 == 'medium clay' &
                       TEXTURE_CODE %in% c('MCS', 'SMC') ~ 'medium clay, sandy, medium',
                     GROUP_2 == 'medium clay' &
                       TEXTURE_CODE %in% c('FSMC', 'MCFS') ~ 'medium clay, sandy, fine',
                     GROUP_2 == 'medium clay' &
                       TEXTURE_CODE %in% c('MCZ', 'ZMC') ~ 'medium clay, silty',
                     GROUP_2 == 'medium clay' &
                       TEXTURE_CODE %in% c('MC') ~ 'medium clay',
                     GROUP_2 == 'medium heavy clay' &
                       TEXTURE_CODE %in% c('KSMHC') ~ 'medium heavy clay, sandy, coarse',
                     GROUP_2 == 'medium heavy clay' &
                       TEXTURE_CODE %in% c('MHCS', 'SMHC') ~ 'medium heavy clay, sandy, medium',
                     GROUP_2 == 'medium heavy clay' &
                       TEXTURE_CODE %in% c('FSMHC', 'MHCFS') ~ 'medium heavy clay, sandy, fine',
                     GROUP_2 == 'medium heavy clay' &
                       TEXTURE_CODE %in% c('ZMHC') ~ 'medium heavy clay, silty',
                     GROUP_2 == 'medium heavy clay' &
                       TEXTURE_CODE %in% c('MHC') ~ 'medium heavy clay',
                     GROUP_2 == 'heavy clay' &
                       TEXTURE_CODE %in% c('KSHC') ~ 'heavy clay, sandy, coarse',
                     GROUP_2 == 'heavy clay' &
                       TEXTURE_CODE %in% c('SHC') ~ 'heavy clay, sandy, medium',
                     GROUP_2 == 'heavy clay' &
                       TEXTURE_CODE %in% c('FSHC') ~ 'heavy clay, sandy, fine',
                     GROUP_2 == 'heavy clay' &
                       TEXTURE_CODE %in% c('ZHC') ~ 'heavy clay, silty',
                     GROUP_2 == 'heavy clay' &
                       TEXTURE_CODE %in% c('HC') ~ 'heavy clay',
                     TRUE ~ NA_character_)) %>% 
  mutate(GROUP_3 = 
           factor(GROUP_3, 
                  levels = 
                    c('gravel', 
                      'sand, coarse', 'sand, medium', 'sand, fine',
                      'loamy sand, coarse', 'loamy sand, medium', 'loamy sand, fine',
                      'clayey sand, coarse', 'clayey sand, medium', 'clayey sand, fine',
                      'peat, coarse', 'peat, medium', 'peat, fine',
                      'mineral peat, coarse', 'mineral peat, medium','mineral peat, fine',
                      'sandy loam, coarse', 'sandy loam, medium', 'sandy loam, fine',
                      'loam, sandy', 'loam, sandy, fine', 'loam', 'silty loam',
                      'sandy clay loam, coarse', 'sandy clay loam, medium', 'sandy clay loam, fine',
                      'clay loam, sandy, coarse', 'clay loam, sandy, medium', 'clay loam, sandy, fine',
                      'clay loam', 'clay loam, silty',  
                      'light clay, sandy, coarse', 'light clay, sandy, medium', 'light clay, sandy, fine',
                      'light clay', 'light clay, silty',
                      'light medium clay, sandy, coarse', 'light medium clay, sandy, medium',
                      'light medium clay, sandy, fine', 'light medium clay', 'light medium clay, silty',
                      'medium clay, sandy, coarse', 'medium clay, sandy, medium',
                      'medium clay, sandy, fine', 'medium clay',  'medium clay, silty',
                      'medium heavy clay, sandy, coarse', 'medium heavy clay, sandy, medium',
                      'medium heavy clay, sandy, fine', 'medium heavy clay', 'medium heavy clay, silty',
                      'heavy clay, sandy, coarse', 'heavy clay, sandy, medium',
                      'heavy clay, sandy, fine', 'heavy clay', 'heavy clay, silty'
                      ),
                  ordered = TRUE)) %>% 
    arrange(GROUP_3)

```

### Group 4

Group 4 rankings account for texture qualifiers, so these have to be re-joined.

In terms of overall 'lightness', the qualifiers induce an internal ranking of 

   * light < sapric < normal < fibric < heavy 
   
(subject to checking, but it makes sense to me...).

```{r grp4}
tc_group4 <- 
  left_join(all_tcs, tc_ranks, by = c('TEXTURE_CODE', 'DESCRIPTION')) %>% 
  mutate(GROUP_4 =
           case_when(TEXTURE_QUALIFIER == '-' & GROUP != 'clay' ~ 
                       paste0('light ', as.character(GROUP_3)),
                     TEXTURE_QUALIFIER == '+' & GROUP != 'clay' ~ 
                       paste0('heavy ', as.character(GROUP_3)),
                     TEXTURE_QUALIFIER == 'A' ~ paste0('sapric ', as.character(GROUP_3)),
                     TEXTURE_QUALIFIER == 'I' ~ paste0('fibric ', as.character(GROUP_3)),
                     TRUE ~ as.character(GROUP_3)),
         # fix where stupid
         GROUP_4 = case_when(GROUP_4 == 'sapric peat, fine' ~ 'peat, fine',
                             GROUP_4 == 'fibric peat, coarse' ~ 'peat, coarse',
                             GROUP_4 == 'fibric peat, medium' ~ 'peat, medium',
                             GROUP_4 == 'heavy gravel' ~ 'gravel',
                             GROUP_4 == 'light gravel' ~ 'gravel',
                             TRUE ~ GROUP_4)) %>% 
  filter(!is.na(TEXTURE_CODE)) %>% 
  mutate(
    GROUP_4 = 
      factor(GROUP_4, 
             levels = 
               c('gravel', 
                 'light sand, coarse','sapric sand, coarse', 'sand, coarse',
                 'fibric sand, coarse', 'heavy sand, coarse',
                 'light sand, medium', 'sapric sand, medium', 'sand, medium',
                 'fibric sand, medium', 'heavy sand, medium',
                 'light sand, fine', 'sand, fine', 'fibric sand, fine', 'heavy sand, fine',
                 'light loamy sand, coarse', 'sapric loamy sand, coarse', 'loamy sand, coarse',
                 'heavy loamy sand, coarse',
                 'light loamy sand, medium', 'sapric loamy sand, medium', 'loamy sand, medium',
                 'fibric loamy sand, medium', 'heavy loamy sand, medium',
                 'light loamy sand, fine', 'sapric loamy sand, fine', 'loamy sand, fine',
                 'fibric loamy sand, fine', 'heavy loamy sand, fine',
                 'light clayey sand, coarse', 'sapric clayey sand, coarse', 'clayey sand, coarse',
                 'fibric clayey sand, coarse', 'heavy clayey sand, coarse',
                 'light clayey sand, medium', 'sapric clayey sand, medium', 'clayey sand, medium',
                 'fibric clayey sand, medium', 'heavy clayey sand, medium',
                 'light clayey sand, fine', 'clayey sand, fine', 'heavy clayey sand, fine',
                 'peat, coarse', 'peat, medium', 'peat, fine',
                 'mineral peat, coarse', 'mineral peat, medium', 'mineral peat, fine',
                 'light sandy loam, coarse', 'sapric sandy loam, coarse', 'sandy loam, coarse',
                 'fibric sandy loam, coarse', 'heavy sandy loam, coarse',
                 'light sandy loam, medium', 'sapric sandy loam, medium', 'sandy loam, medium',
                 'fibric sandy loam, medium', 'heavy sandy loam, medium',
                 'light sandy loam, fine', 'sapric sandy loam, fine', 'sandy loam, fine',
                 'fibric sandy loam, fine','heavy sandy loam, fine',
                 'light loam, sandy', 'loam, sandy', 'heavy loam, sandy',
                 'light loam, sandy, fine', 'loam, sandy, fine', 'fibric loam, sandy, fine',
                 'heavy loam, sandy, fine',
                 'light loam', 'sapric loam', 'loam', 'fibric loam', 'heavy loam',
                 'light sandy clay loam, coarse', 'sapric sandy clay loam, coarse', 'sandy clay loam, coarse',
                 'fibric sandy clay loam, coarse', 'heavy sandy clay loam, coarse',
                 'light sandy clay loam, medium', 'sapric sandy clay loam, medium', 'sandy clay loam, medium',
                 'fibric sandy clay loam, medium', 'heavy sandy clay loam, medium',
                 'light sandy clay loam, fine', 'sapric sandy clay loam, fine', 'sandy clay loam, fine', 
                 'fibric sandy clay loam, fine', 'heavy sandy clay loam, fine',
                 'light silty loam', 'sapric silty loam', 'silty loam', 
                 'fibric silty loam', 'heavy silty loam',
                 'light clay loam, sandy, coarse', 'sapric clay loam, sandy, coarse',
                   'clay loam, sandy, coarse', 'fibric clay loam, sandy, coarse', 
                   'heavy clay loam, sandy, coarse',
                 'light clay loam, sandy, medium', 'sapric clay loam, sandy, medium',
                   'clay loam, sandy, medium', 'fibric clay loam, sandy, medium', 
                   'heavy clay loam, sandy, medium',
                 'light clay loam, sandy, fine', 'sapric clay loam, sandy, fine',
                   'clay loam, sandy, fine', 'fibric clay loam, sandy, fine', 
                   'heavy clay loam, sandy, fine',
                 'light clay loam', 'sapric clay loam', 'clay loam', 'fibric clay loam',
                   'heavy clay loam',
                 'light clay loam, silty', 'sapric clay loam, silty',
                   'clay loam, silty', 'fibric clay loam, silty', 'heavy clay loam, silty',
                 'sapric light clay, sandy, coarse', 'light clay, sandy, coarse', 
                   'fibric light clay, sandy, coarse',
                 'sapric light clay, sandy, medium', 'light clay, sandy, medium', 
                   'fibric light clay, sandy, medium',
                 'sapric light clay, sandy, fine', 'light clay, sandy, fine', 
                   'fibric light clay, sandy, fine',
                 'sapric light clay', 'light clay', 'fibric light clay',
                 'sapric light clay, silty', 'light clay, silty', 'fibric light clay, silty',
                 'light medium clay, sandy, coarse',
                 'sapric light medium clay, sandy, medium', 'light medium clay, sandy, medium',
                 'fibric light medium clay, sandy, medium',
                 'sapric light medium clay, sandy, fine', 'light medium clay, sandy, fine',
                 'fibric light medium clay, sandy, fine',
                 'sapric light medium clay', 'light medium clay', 'fibric light medium clay',
                 'sapric light medium clay, silty', 'light medium clay, silty',
                 'fibric light medium clay, silty',
                 'medium clay, sandy, coarse', 'fibric medium clay, sandy, coarse',
                 'sapric medium clay, sandy, medium', 'medium clay, sandy, medium',
                 'fibric medium clay, sandy, medium',
                 'medium clay, sandy, fine', 'fibric medium clay, sandy, fine',
                 'sapric medium clay', 'medium clay', 'fibric medium clay',
                 'sapric medium clay, silty', 'medium clay, silty',
                 'fibric medium clay, silty',
                 'medium heavy clay, sandy, coarse', 'medium heavy clay, sandy, medium',
                 'medium heavy clay, sandy, fine', 'sapric medium heavy clay', 'medium heavy clay', 'fibric medium heavy clay',
                 'medium heavy clay, silty', 'fibric medium heavy clay, silty',
                 'heavy clay, sandy, coarse', 'heavy clay, sandy, medium', 'heavy clay, sandy, fine',
                 'sapric heavy clay', 'heavy clay', 'fibric heavy clay',
                 'heavy clay, silty', 'fibric heavy clay, silty'),
             ordered = TRUE),
    FINE_RANK = as.integer(GROUP_4), 
    TQ_FCT = case_when(is.na(TEXTURE_QUALIFIER) ~ 'N',
                       TRUE ~ TEXTURE_QUALIFIER),
    TQ_FCT = factor(TQ_FCT, levels = c('-', 'A', 'N', 'I', '+'), ordered = TRUE),
    TQ_FCT = as.integer(TQ_FCT)) %>% 
  arrange(GROUP_4, TQ_FCT) %>% 
  mutate(FINAL_RANK = seq(nrow(.)))

all_tc_ranked <- all_tcs %>% 
  filter(!is.na(TEXTURE_CODE)) %>% 
  dplyr::select(TEXTURE, TEXTURE_CODE, TEXTURE_QUALIFIER, DESCRIPTION) %>% 
  left_join(tc_group4, by = c('TEXTURE', 'TEXTURE_CODE', 'TEXTURE_QUALIFIER', 'DESCRIPTION')) %>% 
  filter(!is.na(TEXTURE_CODE)) %>% 
  dplyr::select(TEXTURE, TEXTURE_CODE, TEXTURE_QUALIFIER, DESCRIPTION,
                GROUP, GROUP_2, GROUP_3, GROUP_4, N_CODES, FINAL_RANK) %>% 
  rename(GROUP_1 = GROUP) %>% 
  arrange(FINAL_RANK) 

write_csv(all_tc_ranked, file.path(getwd(), 'helpers', 'SALI_TEXTURES_RANKED_2.csv'))

```

***

