---
title: "HELPER - Munsell Colours to Hex values"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hold', fig.height = 6,
                      fig.align = 'center', warning = FALSE, message = FALSE,
                      eval = TRUE)
```

Munsell soil colour notations need web-friendly hex equivalents for devs and cartos.

```{r pkgs, results = 'hide'}
options(stringsAsFactors = FALSE, scipen = 999)
library(tidyverse)       # general data manipulation and display
library(munsellinterpol) # munsell translation

rgb2hex <- function(r,g,b) rgb(r, g, b, maxColorValue = 255)
```

To generate hex colours for the munsell codes used in sit_horizon_colours and sit_mottles,

```{r longway}
today <- '20191028' 
sit_datafile <- file.path(getwd(), 'SALI_DATA', 
                          paste0('SALI_SIT-data_', today, '.rds'))
dec_datafile <- file.path(getwd(), 'SALI_DATA', 
                          paste0('SALI_SIT-data_decodes_', today, '.rds'))

sit_data <- readRDS(sit_datafile)
decodes <- readRDS(dec_datafile)

hor_cols <- sit_data[['HCO']] %>% 
  dplyr::select(COLOUR_BOOK, COLOUR_CODE) %>% 
  distinct() 

mott_cols <- sit_data[['HMT']] %>% 
  dplyr::filter(COLOUR_BOOK != 'Y') %>% 
  dplyr::select(COLOUR_BOOK, COLOUR_CODE) %>% 
  distinct()

all_cols <- rbind(hor_cols, mott_cols) %>% 
  distinct() %>% 
  left_join(., decodes[['COL']], by = c('COLOUR_BOOK', 'COLOUR_CODE')) %>% 
  dplyr::select(COLOUR_BOOK, COLOUR_CODE, HUE, VALUE, CHROMA) %>% 
  mutate(COL_FORMATTED = paste0(HUE, ' ', VALUE, '/', CHROMA)) %>% 
  filter(!is.na(COLOUR_CODE))

all_cols_translated <- lapply(all_cols$COL_FORMATTED, MunsellToRGB) %>% 
  map_dfr(., function(x) {
    data.frame("COLOURCODE" = x$SAMPLE_NAME,
               "R" = round(x$RGB[1]),
               "G" = round(x$RGB[2]),
               "B" = round(x$RGB[3])
               ) 
  }) %>% 
  mutate(HEX = rgb2hex(R,G,B)) %>% 
  cbind(all_cols, .) %>% 
  select(-COL_FORMATTED, -COLOURCODE) %>% 
  arrange(desc(COLOUR_BOOK), COLOUR_CODE)

write_csv(all_cols_translated, 
          file.path(getwd(), 'helpers', 'translated_colours.csv'))
```

A minority of mottles are described with reference to Munsell codes - most are described using the simpler Yellow Book colours. May as well supply a hex value here too, with reference to common values seen in the field. Nominating some fairly arbitrary hex values for Yellow Book mottle colours as:

  * R Red    - 10R48  - #9A4832
  * O Orange - 5YR78  - #E69D64
  * B Brown  - 10YR46 - #7F5921
  * Y Yellow - 5Y88   - #E2C75A
  * G Grey   - 10YR61 - #9B9288
  * D Dark   - 10YR31 - #4F463E
  * L Gley   - 10GY51 - #81786F
  * P Pale   - 10YR81 - #CFC7BB
  
This might help with e.g. code-generated soil profile diagrams, but users should be aware that the YB colours conceal a fair bit of variation, particularly the dark/brown/pale codes.

