---
title: "Morphology WFS - Option 3"
output: html_document
author: "Lauren O'Brien"
date: `r Sys.Date()`
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'hold', fig.height = 6,
                      fig.align = 'center', warning = FALSE, message = FALSE,
                      eval = FALSE)
```

### Background

Experiments with extracting and formatting Queensland soils morphology data for WFS publication. Primary driver is the TERN SGLA II modelling (and particularly interface with the soil data federator https://github.com/RossDSearle/SoilDataFederatoR/) but the service should be useful to a wider range of clients as well.

```{r pkgs, results = 'hide'}
options(stringsAsFactors = FALSE, scipen = 999)
library(sf)            # vector data IO and processing
library(DBI)           # database interface
library(ROracle)       # Oracle database drivers
library(getPass)       # password management
library(units)         # for working with unit-attributed data
library(tidyverse)     # general data manipulation and display

Sys.setenv(
  'TNS_ADMIN'   = file.path('C:', 'Users', Sys.getenv('USERNAME'), 'Oracle'),
  'ORACLE_HOME' = 'C:/Program Files/Oracle Instant Client'
  )
SALI <- dbConnect(dbDriver('Oracle'),
                  username = 'obrienle',
                  password = getPass(),
                  dbname   = 'sispda',
                  host     = 'ex02client01',
                  port     = 1521)
```

Earlier work - still mucking around in here 

## Structure

  * Layer 1 - Sites summary data (to observation level plus calculated values)
  * Layer 2 - Soil Classification (all data)
  * Layer 3 - Disturbance, Microrelief, Erosion and Surface Condition
  * Layer 4 - Surface Coarse Fragments and Rock Outcrop
  * Layer 5 - Horizon definitions (name, thickness, colour, texture etc plus some summaries tba)
  * Layer 6 - Horizon colours
  * Layer 7 - Horizon mottles 
  * Layer 8 - Horizon structures
  * Layer 9 - Horizon coarse fragments, segregations and pans
  * Layer 10 - Horizon cutans and fabrics
  * Layer 11 - Horizon cracks, macropores and roots
  * Layer 12 - Field tests
  * Layer 13 - Samples (maybe with cross link to soil chem API)

Load data snapshot from 'DATA_sali-extract.Rmd'.

```{r getdat}
today <- '20191212' 
sit_datafile <- file.path(getwd(), 'SALI_DATA', 
                          paste0('SALI_SIT-data_', today, '.rds'))
sit_data <- readRDS(sit_datafile)
sit_data[['LAB']] <- NULL # out of scope still
```

### Layer 1

```{r layer1}
sitobs_base <- 
    dbGetQuery(SALI, 
               "select project_code, site_id, obs_no,
                 obs_date, site_type, obs_type,
                 longitude, latitude, loc_accuracy, loc_meas_method,
                 runoff, permeability, drainage, lith_code, land_use_code,
                 slope_eval_method, slope_pct, slope_class, slope_morph_type, 
                 elem_inc_slope, elem_type_code,
                 modal_slope, relief_class, rel_mod_slope_class, patt_type_code
               from reg_projects
               join sit_sites using (project_code)
               join sit_observations using (project_code, site_id)
               join sit_locations using (project_code, site_id, obs_no)
               where avail_status = 'E'
               and site_qc_status = 'P'
               and datum = 3
               order by project_code, site_id, obs_no") %>% 
    dplyr::filter(LONGITUDE > 1) # >:-(

 # add in summary stuff from L1
```

### Layer 2 - Classifications

Ditching `TECH_REF_CODE` because the contents don't currently make any sense. ASC level data is also slightly stuffed atm. Renaming some of the ASC fields for greater consistency and easier querying.

```{r layer2}
loc <- sit_data[['LOC']] %>%  
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, LONGITUDE, LATITUDE)

layer_2 <- sit_data[['CLA']] %>% 
  rename(ASC_SORD = SUBORD_ASC_CODE,
         ASC_GGRP = GREAT_GROUP_ASC_CODE,
         ASC_SGRP = SUBGROUP_ASC_CODE,
         GSG_1    = EXHIBIT_GSG_CODE,
         GSG_2    = AFFINITY_GSG_CODE
         ) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, SOIL_CLASS_NO, BURIED_FLAG,
                ASC_CONFIDENCE, ASC_ORD, ASC_SORD, ASC_GGRP, ASC_SGRP,
                ASC_FAM1, ASC_FAM2, ASC_FAM3, ASC_FAM4, ASC_FAM5, PPF,
                GSG_1, GSG_2) %>% 
  left_join(., loc,
            by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>%
  dplyr::arrange(PROJECT_CODE, SITE_ID, OBS_NO, SOIL_CLASS_NO) %>% 
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283)
  
# e.g. 
plot(dplyr::filter(layer_2, PROJECT_CODE == 'BAB')['ASC_ORD'], pch = 20)

```

### Layer 3 - Disturbance, Microrelief, Erosion and Surface Condition

These are all soil surface parameters where multiple entries are possible.

Microrelief is massaged a bit here for clarity - the yellow book codes are somewhat poorly structured.

```{r layer3}
dis <- sit_data[['DIS']] %>% 
  mutate(PARAM = 'DISTURBANCE') %>% 
  rename(PARAM_NO = DISTURB_NO) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, PARAM, PARAM_NO, everything())

ero <- sit_data[['ERO']] %>% 
  mutate(PARAM = 'EROSION') %>% 
  rename(PARAM_NO        = ERO_NO,
         ERO_GULLY_DEPTH = GULLY_DEPTH) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, PARAM, PARAM_NO, everything())

mic <- sit_data[['MIC']] %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  mutate(PARAM = 'MICRORELIEF',
         MICRO_GROUP = 
           case_when(MICRO_TYPE == 'Z'                 ~ 'Z', # None
                     MICRO_TYPE %in% 
                       c('C', 'N', 'L', 'A', 'M', 'G') ~ 'G', # Gilgai
                     MICRO_TYPE %in% c('D', 'W')       ~ 'H', # Hummocky
                     MICRO_TYPE %in% 
                       c('U', 'K', 'I', 'S', 'R', 'T',
                         'P', 'H', 'O')                ~ 'O', # Other,
                     MICRO_TYPE == 'B'                 ~ 'B', # Biotic
                     TRUE ~ NA_character_),
         # MICRO_TYPE usually '-' or 'Z' or 'O' which isn't useful
         # also a fair few mistakes like 'N' with 'O'
         MICRO_GROUP = ifelse(is.na(BIOTIC_AGENT), MICRO_GROUP, 'B')) %>%
  rename(PARAM_NO         = MICRO_NO, 
         MIC_PROP_GILGAI  = PROPORTION_GILGAI,
         MIC_VI           = VERTICAL_INTERVAL,
         MIC_HI           = HORIZ_INTERVAL) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, PARAM, PARAM_NO, MICRO_GROUP,
                MICRO_TYPE, BIOTIC_AGENT, BIOTIC_COMP, everything())

scn <- sit_data[['SCN']] %>% 
  mutate(PARAM = 'SURF_COND') %>% 
  rename(PARAM_NO = SURF_COND_NO, SURF_COND_STATUS = STATUS) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, PARAM, PARAM_NO, everything())

layer_3 <- reduce(list(dis, scn, mic, ero),
                  function(x, y) {
                    full_join(x, y, by = intersect(names(x), names(y)))
                    }
                  ) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, PARAM, PARAM_NO, everything()) %>% 
  left_join(., loc,
            by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>%
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283)
                
# e.g.
plot(dplyr::filter(layer_3, PROJECT_CODE == 'BAB')['DISTURB_TYPE'], pch = 20)

```

### Layer 4 - Surface Coarse Fragments and Rock Outcrop

NB review abundance column as the decodes aren't quite the same...

```{r layer4}

scf <- sit_data[['SCF']] %>% 
  mutate(PARAM = 'SURF_FRAGS') %>% 
  rename(PARAM_NO = SURF_FRAG_NO) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, PARAM, PARAM_NO, everything())

otc <- sit_data[['OTC']] %>% 
  mutate(PARAM = 'OUTCROP') %>% 
  rename(PARAM_NO = OUTCROP_NO) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, PARAM, PARAM_NO, everything())

layer_4 <- reduce(list(scf, otc),
                  function(x, y) {
                    full_join(x, y, by = intersect(names(x), names(y)))
                    }
                  ) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, PARAM, PARAM_NO,
                ABUNDANCE, LITH_CODE, ALTERATION, 
                everything()) %>% 
  left_join(., loc,
            by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>%
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283) 

# e.g.
plot(dplyr::filter(layer_4, PROJECT_CODE == 'BAB')['LITH_CODE'], pch = 20)
# that's actually quite interesting!
```

### Layer 5 - Horizon definitions

NB Consider folding in a concat col of suffixes

```{r layer5}

layer_5 <- sit_data[['HOR']] %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO,
                UPPER_DEPTH, LOWER_DEPTH, 
                HORIZON_NAME, TEXTURE, PERMEABILITY, DRAINAGE, 
                DESIGN_NUM_PREFIX, DESIGN_MASTER, DESIGN_SUBDIV,
                everything()) %>% 
  left_join(., loc, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283) 

# e.g.
plot(dplyr::filter(layer_5, PROJECT_CODE == 'BAB' &
                     HORIZON_NO == 1)['TEXTURE_CODE'], pch = 20)

```

### Layer 6 - Horizon colours

Adding in hex values and also splitting up munsell to HVC components. Note Japanese and Munsell systems match where the codes are the same, but some codes were not available in one book vs the other.

```{r layer6}
sit_colours <- dbGetQuery(SALI, 
                          "select colour_book, colour_code, hue, value, chroma, 
                           rgb_value as hex_value
                           from sit_colours col
                           join sit_colour_convert ccv 
                           on ccv.sali_colour = col.colour_code")

hdepths <- sit_data[['HOR']] %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, 
                UPPER_DEPTH, LOWER_DEPTH)

layer_6 <- sit_data[['COL']] %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  left_join(., sit_colours, by = c('COLOUR_BOOK', 'COLOUR_CODE')) %>% 
  left_join(., loc, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  left_join(., hdepths, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO',
                               'HORIZON_NO')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, 
                UPPER_DEPTH, LOWER_DEPTH, everything()) %>% 
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283) 

# e.g.
plot(dplyr::filter(layer_6, PROJECT_CODE == 'BAB' &
                     HORIZON_NO == 1 & HOR_COL_NO == 1)['HUE'], pch = 20)
   
```

### Layer 7 - Horizon mottles 

A minority of mottles are described with reference to Munsell codes - most are described using the simpler Yellow Book colours. May as well supply a hex value here too, with reference to common values seen in the field. Nominating some fairly arbitrary hex values for Yellow Book mottle colours as:

  * R Red    - 10R48  - #9A4832
  * O Orange - 5YR78  - #E69D64
  * B Brown  - 10YR46 - #7F5921
  * Y Yellow - 5Y88   - #E2C75A
  * G Grey   - 10YR61 - #9B9288
  * D Dark   - 10YR31 - #4F463E
  * L Gley   - 10GY51 - #81786F
  * P Pale   - 10YR81 - #CFC7BB
  
This might help with e.g. code-generated soil profile diagrams, but users should be aware that the YB colours conceal a fair bit of variation, particularly the dark/brown/pale codes.

```{r layer7}

layer_7 <- sit_data[['MOT']] %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  rename(MOTT_ABUND = ABUNDANCE,
         MOTT_BOUND_DISTINCT = BOUND_DISTINCT) %>% 
  left_join(., sit_colours, by = c('COLOUR_BOOK', 'COLOUR_CODE')) %>% 
  mutate(HEX_VALUE = 
           case_when(COLOUR_BOOK == 'Y' & COLOUR_CODE == 'R' ~ '994730',
                     COLOUR_BOOK == 'Y' & COLOUR_CODE == 'O' ~ 'E89E63',
                     COLOUR_BOOK == 'Y' & COLOUR_CODE == 'B' ~ '805921',
                     COLOUR_BOOK == 'Y' & COLOUR_CODE == 'Y' ~ 'E1C85A',
                     COLOUR_BOOK == 'Y' & COLOUR_CODE == 'G' ~ '9E9187',
                     COLOUR_BOOK == 'Y' & COLOUR_CODE == 'D' ~ '4F453D',
                     COLOUR_BOOK == 'Y' & COLOUR_CODE == 'L' ~ '737D73',
                     COLOUR_BOOK == 'Y' & COLOUR_CODE == 'P' ~ 'D1C7BD',
                     TRUE ~ HEX_VALUE)) %>% 
  left_join(., hdepths, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO',
                               'HORIZON_NO')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, UPPER_DEPTH, 
                LOWER_DEPTH, MOTT_NO,
                MOTT_TYPE, MOTT_ABUND, MOTT_SIZE, CONTRAST, COLOUR_CODE, 
                MOISTURE_STAT, MOTT_BOUND_DISTINCT, everything()) %>% 
  left_join(., loc, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283) 
 

```

### Layer 8 - Horizon structures

```{r layer8}

layer_8 <- sit_data[['STR']] %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  left_join(., hdepths, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO',
                               'HORIZON_NO')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, UPPER_DEPTH, 
                LOWER_DEPTH, STRUCT_NO, everything()) %>% 
  left_join(., loc, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283)

# e.g.
plot(dplyr::filter(layer_8, PROJECT_CODE == 'BAB' &
                     HORIZON_NO == 1 & STRUCT_NO == 1)['PEDALITY_GRADE'], 
     pch = 20)

```

### Layer 9 - Horizon coarse fragments, segregations and pans

Lumping the lumps together >.>

```{r layer9}

hcf <- sit_data[['CFS']] %>% 
  mutate(PARAM = 'COARSE_FRAGS') %>% 
  rename(PARAM_NO = COARSE_FRAG_NO) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, PARAM, PARAM_NO,
                ABUNDANCE, COARSE_FRAG_SIZE, SHAPE, LITH_CODE, everything())

seg <- sit_data[['SEG']] %>% 
  mutate(PARAM = 'SEGREGATIONS') %>% 
  rename(PARAM_NO = SEG_NO,
         SEG_NATURE = NATURE) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, PARAM, PARAM_NO,
                everything())

pan <- sit_data[['PAN']] %>% 
  mutate(PARAM = 'PANS') %>% 
  rename(PARAM_NO = PAN_NO,
         PAN_CEMENT = CEMENTATION,
         PAN_CONTIN = CONTINUITY,
         PAN_STRUCTURE = STRUCTURE) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, PARAM, PARAM_NO,
                everything())

layer_9 <- reduce(list(hcf, seg, pan),
                  function(x, y) {
                    full_join(x, y, by = intersect(names(x), names(y)))
                    }
                  ) %>% 
  left_join(., hdepths, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO',
                               'HORIZON_NO')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, UPPER_DEPTH, 
                LOWER_DEPTH, PARAM, PARAM_NO, ABUNDANCE, LITH_CODE, 
                ALTERATION, everything()) %>%
  left_join(., loc, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283) 

# e.g.
plot(dplyr::filter(layer_9, PROJECT_CODE == 'BAB' &
                     HORIZON_NO == 1 & PARAM_NO == 1)['ABUNDANCE'], 
     pch = 20)

```

### Layer 10 - Horizon cutans and fabrics

Ped-surface features

```{r layer_10}

cut <- sit_data[['CUT']] %>% 
  mutate(PARAM = 'CUTANS') %>% 
  rename(PARAM_NO = CUTAN_NO) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, PARAM, PARAM_NO,
                everything()) 

fab <- sit_data[['FAB']] %>% 
  mutate(PARAM = 'FABRICS') %>% 
  rename(PARAM_NO = FAB_NO) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, PARAM, PARAM_NO,
                everything())

layer_10 <- reduce(list(cut, fab),
                  function(x, y) {
                    full_join(x, y, by = intersect(names(x), names(y)))
                    }
                  ) %>% 
  left_join(., hdepths, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO',
                               'HORIZON_NO')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, UPPER_DEPTH,
                LOWER_DEPTH, PARAM, PARAM_NO, everything()) %>%
  left_join(., loc, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283) 

# e.g.
plot(dplyr::filter(layer_10, PROJECT_CODE == 'BAB' &
                     HORIZON_NO == 1 & PARAM_NO == 1)['CUTAN_TYPE'], 
     pch = 20)

```

### Layer 11 - Horizon cracks, macropores and roots

The first two are under 'Voids' in the YB. All three facilitate water entry into the soil.

```{r layer11}

crk <- sit_data[['CRK']] %>% 
  mutate(PARAM = 'CRACKS') %>% 
  rename(PARAM_NO = CRACK_NO,
         CRACK_WIDTH = WIDTH) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, PARAM, PARAM_NO,
                everything())

por <- sit_data[['POR']] %>% 
  mutate(PARAM = 'MACROPORES') %>% 
  rename(PARAM_NO = MACRO_NO,
         PORE_DIAMETER = DIAMETER) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, PARAM, PARAM_NO,
                everything())

rts <- sit_data[['RTS']] %>% 
  mutate(PARAM = 'ROOTS') %>% 
  rename(PARAM_NO = ROOT_NO) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, PARAM, PARAM_NO,
                everything())
layer_11 <- reduce(list(crk, por, rts),
                  function(x, y) {
                    full_join(x, y, by = intersect(names(x), names(y)))
                    }
                  ) %>% 
  left_join(., hdepths, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO',
                               'HORIZON_NO')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, UPPER_DEPTH,
                LOWER_DEPTH, PARAM, PARAM_NO, everything()) %>%
  left_join(., loc, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283) 

```

### Layer 12 -  Field Tests

Maybe I should have just pulled them out as one table!

```{r layer12}

layer_12 <- do.call('rbind', list(sit_data[['FT_PH']], sit_data[['FT_EC']],
                                  sit_data[['FT_DS']], sit_data[['FT_SL']],
                                  sit_data[['FT_FZ']], sit_data[['FT_H2']])) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) %>% 
  left_join(., hdepths, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO',
                               'HORIZON_NO')) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, UPPER_DEPTH,
                LOWER_DEPTH, TEST_NO, everything()) %>% 
  rename(TEST_DEPTH = DEPTH) %>% 
  dplyr::arrange(PROJECT_CODE, SITE_ID, OBS_NO, HORIZON_NO, TEST_NO, DEPTH) %>%
  left_join(., loc, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283) 

# e.g.
sph <- dplyr::filter(layer_12, PROJECT_CODE == 'BAB' &
                     HORIZON_NO == 4 & TEST_TYPE == 'PH') %>% 
       mutate(VALUE = as.numeric(VALUE))
plot(sph['VALUE'],  pch = 20)

```

### Layer 13 - Samples

TODO List available analytes (in general terms e.g. 'has pH, EC, Cations' etc)

Provide cross link to soil-chem API at sample level??? No webpage interface tho... would return JSON.

Query form would be 
```{r getlab, eval = FALSE} 
chem_api <- "https://soil-chem.information.qld.gov.au/odata/"
results_endpoint <- "SiteLabMethodResults?"
filter <- paste0("%24filter=ProjectCode%20eq%20'", project_code,
                 "'%20and%20SiteId%20eq%20", site_id,
                 "%20and%20ObservationNumber%20eq%20", obs_no,
                 "%20and%20SampleNumber%20eq%20", sample_no)
qry <- paste0(chem_api, results_endpoint, filter)
```

```{r layer13}
## mapped lab codes to general groups manually beforehand
#grps <- readxl::read_xlsx(file.path(getwd(), 'helpers',
#                                    'SALI_Lab_Method_Groupings.xlsx'),
#                          sheet = 1)
#
#labstuff <- dbGetQuery(SALI,
#                       "select distinct project_code, site_id, obs_no, 
#                          sample_no, lab_meth_code 
#                        from sit_lab_results 
#                        order by project_code, site_id, obs_no, 
#                          sample_no, lab_meth_code") %>% 
#  left_join(., grps, by = 'LAB_METH_CODE') %>% 
#  select(-LAB_METH_CODE) %>% 
#  group_by(PROJECT_CODE, SITE_ID, OBS_NO, SAMPLE_NO) %>% 
#  summarise(MEAS_TYPES = toString(sort(unique(MEASUREMENT_TYPE))),
#            ANALYTICAL_GROUPS = toString(sort(unique(ANALYTICAL_GROUP))))
# hmmm!

# for now,  
layer_13 <- sit_data[['SAM']] %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT'), -SUBMISS_NOTES) %>% 
  left_join(., loc, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  filter(!is.na(LONGITUDE)) %>% 
  st_as_sf(., coords = c('LONGITUDE', 'LATITUDE'), crs = 4283) 


```

## Assemble service

Will be publishing via SIR so may as well write to FGDB as well as RDS

```{r esritime}

SERVICE_B <- list(layer_1, layer_2, layer_3, layer_4, layer_5, layer_6, layer_7, layer_8, layer_9, layer_10, layer_11, layer_12, layer_13)

names(SERVICE_B) <- paste0("LAYER_", seq(13))

saveRDS(SERVICE_B, file.path(getwd(), 'CLEAN_DATA', 'SERVICE_B.rds'))

# NFI how I'm going to name these in line with the SIR convention - will probs
# have to fall back on QLD_SOIL_SITES_SERVICE_LYR1_P etc blerg




```
