---
title: "Morphology WFS - Option 2"
output: html_document
author: "Lauren O'Brien"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hold', fig.height = 6,
                      fig.align = 'center', warning = FALSE, message = FALSE,
                      eval = TRUE)
```

### Background

Experiments with extracting and formatting Queensland soils morphology data for WFS publication. Primary driver is the TERN SGLA II modelling (and particularly interface with the soil data federator https://github.com/RossDSearle/SoilDataFederatoR/) but the service should be useful to a wider range of clients as well.

```{r pkgs, results = 'hide'}
options(stringsAsFactors = FALSE, scipen = 999)
library(sf)            # vector data IO and processing
library(units)         # for working with unit-attributed data
library(tidyverse)     # general data manipulation and display
```

### 'Flat' conversion

This option emphasises minimising the number of layers in the service to align with the 'join levels' of Site, Observation, Horizon, Veg Community. Keeping samples and notes separate too.

```{r getdat}
today <- '20191101' 
sit_datafile <- file.path(getwd(), 'SALI_DATA', 
                          paste0('SALI_SIT-data_', today, '.rds'))
decode_datafile <- file.path(getwd(), 'SALI_DATA', 
                             paste0('SALI_SIT-data_decodes_', today, '.rds'))

sit_data <- readRDS(sit_datafile)
decodes  <- readRDS(decode_datafile)
```

Pros:

  * Simpler layer model
  * Less repeated geometry data

Cons 

  * Smashing all the tables together neatly could be tricky - will need to alias some column names
  * More repeated data from table joins, more NULL data
  * Lose the ability to return useful created and updated dates 
  * Display could be odd??

#### Sites Layer

Requires reconciling tables SIT, SLU, SMP, SGE, SGP and spatialising using Obs 1.

```{r siteslyr}
site_data <- sit_data[1:5]
site_data <- lapply(site_data, function(x) {
  dplyr::select(x, -matches('CREAT'), -matches('UPDAT')) 
  })

# context Proj code and context sid not in use on these projects so
site_data[['SIT']] <- 
  dplyr::select(site_data[['SIT']], -CONTEXT_PROJECT_CODE, -CONTEXT_SITE_ID)

# duplicate column names in geomorph tables need aliasing 
site_data[['SGE']] <- site_data[['SGE']] %>% 
  rename(ELEM_GEO_AGENT = GEO_AGENT, ELEM_GEO_MODE = GEO_MODE)

site_data[['SGP']] <- site_data[['SGP']] %>% 
  rename(PATT_GEO_AGENT = GEO_AGENT, PATT_GEO_MODE = GEO_MODE)

sites_layer <- reduce(site_data, full_join)

# spatialise
locs <- sit_data[['OLC']] %>% 
  dplyr::filter(OBS_NO == 1) %>% 
  st_as_sf(coords = c('LONGITUDE', 'LATITUDE'), crs = 4283) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID)

sites_layer <- right_join(locs, sites_layer, by = c('PROJECT_CODE', 'SITE_ID'))

names(sites_layer)

# e.g.
demo <- dplyr::filter(sites_layer, PROJECT_CODE == 'GCL') %>% 
  dplyr::filter(!is.na(SLOPE_CLASS)) %>% 
  mutate(SC = factor(SLOPE_CLASS, 
                     levels = c('LE', 'VG', 'GE', 'MO', 'ST', 'VS', 'CL'),
                     labels = c('Level', 'Very Gentle', 'Gentle', 'Moderate',
                                'Steep', 'Very Steep', 'Cliffed'), 
                     ordered = TRUE))

ggplot() +
  geom_sf(data = demo, aes(colour = SC), pch = 16, size = 2,
          show.legend = 'point') +
  scale_colour_viridis_d() +
  ggtitle("Slope Class", subtitle = "SALI Project GCL") +
  labs(colour = 'Slope Class') +
  theme_minimal()

```

#### Observations layer

Requires reconciling tables OBS, OLC, OCL, ODS, OSC, OMC, OER, OCF, OTC, OSM

```{r obslyr}
obs_data <- sit_data[7:16]
obs_data <- lapply(obs_data, function(x) {
  dplyr::select(x, -matches('CREAT'), -matches('UPDAT')) 
  })

# redundant redundant
obs_data[['OLC']] <- NULL

# going to separate obs level veg data out 
obs_data[['OBS']] <- obs_data[['OBS']] %>% 
  dplyr::select(-FOREST_TYPE, -matches('RAINFOR'), -SCLEROPHYLL)

obs_layer <- reduce(obs_data, full_join) # still not quite right?? check

locs <- sit_data[['OLC']] %>% 
  st_as_sf(coords = c('LONGITUDE', 'LATITUDE'), crs = 4283) %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO)

obs_layer <- 
  right_join(locs, obs_layer, by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO'))

names(obs_layer)

# e.g.
demo <- dplyr::filter(obs_layer, PROJECT_CODE == 'GCL') %>% 
  dplyr::filter(!is.na(DRAINAGE)) %>% 
  mutate(DG = factor(DRAINAGE, 
                     levels = c(1, 2, 3, 4, 5, 6),
                     labels = c('Very Poor', 'Poor', 'Imperfect', 'Moderate', 
                                'Well', 'Very Well'), 
                     ordered = TRUE))

ggplot() +
  geom_sf(data = demo, aes(colour = DG), pch = 16, size = 2,
          show.legend = 'point') +
  scale_colour_viridis_d() +
  ggtitle("Drainage Class", subtitle = "SALI Project GCL") +
  labs(colour = 'Drainage Class') +
  theme_minimal()
```

#### Horizons layer

Requires reconciling tables HOR, HSX, HCO, HMT, HCF, HST, HSG, HFB, HCU, HCR, HMP, HGT, HPA, HRT, FTS

Skipping HSX as all the data actually appears in HOR$HORIZON_NAME

```{r horlyr}
hor_data <- sit_data[20:34]
hor_data[['HSX']] <- NULL
hor_data <- lapply(hor_data, function(x) {
  dplyr::select(x, -matches('CREAT'), -matches('UPDAT')) 
  })

# also need hex colours
hex_cols <- read_csv(file.path(getwd(), 'helpers', 'translated_colours.csv')) %>% 
  dplyr::select(COLOUR_BOOK, COLOUR_CODE, HEX)

# so much aliasing augh
hor_data[['HOR']] <- hor_data[['HOR']] %>% 
  rename(HOR_BOUND_DISTINCT = BOUND_DISTINCT, HOR_BOUND_SHAPE = BOUND_SHAPE)

hor_data[['HCO']] <- hor_data[['HCO']] %>% 
  left_join(hex_cols, by = c('COLOUR_BOOK', 'COLOUR_CODE')) %>% 
  rename(HOR_COLOUR_BOOK = COLOUR_BOOK, HOR_COLOUR_CODE = COLOUR_CODE,
         HOR_MOISTURE_STAT = MOISTURE_STAT, HOR_HEXCODE = HEX)

hor_data[['HMT']] <- hor_data[['HMT']] %>%  
  left_join(hex_cols, by = c('COLOUR_BOOK', 'COLOUR_CODE')) %>% 
  rename(MOTT_COLOUR_BOOK = COLOUR_BOOK, MOTT_COLOUR_CODE = COLOUR_CODE, 
         MOTT_MOISTURE_STAT = MOISTURE_STAT, MOTT_HEXCODE = HEX,
         MOTT_CONTRAST = CONTRAST, MOTT_ABUNDANCE = ABUNDANCE, 
         MOTT_BOUND_DISTINCT = BOUND_DISTINCT)

hor_data[['HCF']] <- hor_data[['HCF']] %>%
  rename(COARSE_FRAG_ABUNDANCE = ABUNDANCE, COARSE_FRAG_LITH_CODE = LITH_CODE,
         COARSE_FRAG_SHAPE = SHAPE, COARSE_FRAG_STRENGTH = STRENGTH,
         COARSE_FRAG_DISTRIBUTION = DISTRIBUTION, 
         COARSE_FRAG_ALTERATION = ALTERATION )

# structures is ok

hor_data[['HSG']] <- hor_data[['HSG']] %>%
  rename(SEG_ABUNDANCE = ABUNDANCE, SEG_NATURE = NATURE, 
         SEG_STRENGTH = STRENGTH, SEG_MAGNETIC_ATTRIB = MAGNETIC_ATTRIB)

hor_data[['HFB']] <- hor_data[['HFB']] %>%
  rename(FAB_ABUNDANCE = ABUNDANCE)

hor_data[['HCU']] <- hor_data[['HCU']] %>%
  rename(CUTAN_ABUNDANCE = ABUNDANCE, CUTAN_DISTINCTNESS = DISTINCTNESS)

hor_data[['HCR']] <- hor_data[['HCR']] %>%
  rename(CRACK_WIDTH = WIDTH)

hor_data[['HMP']] <- hor_data[['HMP']] %>%
  rename(MACRO_ABUNDANCE = ABUNDANCE, MACRO_DIAMETER = DIAMETER)

hor_data[['HGT']] <- hor_data[['HGT']] %>%
  rename(STREN_CLASS = CLASS, STREN_WATER_STAT = WATER_STAT)

hor_data[['HPA']] <- hor_data[['HPA']] %>%
  rename(PAN_CEMENTATION = CEMENTATION, PAN_CONTINUITY = CONTINUITY, 
         PAN_STRUCTURE = STRUCTURE)

hor_data[['HRT']] <- hor_data[['HRT']] %>%
  rename(ROOT_ABUNDANCE = ABUNDANCE)

hor_data[['FTS']] <- hor_data[['FTS']] %>%
  rename(FIELD_TEST_NO = TEST_NO, FIELD_TEST_TYPE = TEST_TYPE, 
         FIELD_TEST_DEPTH = DEPTH, FIELD_TEST_VALUE = VALUE, 
         FIELD_TEST_METHOD = METHOD)

horizon_layer <- reduce(hor_data, full_join) %>% 
  right_join(locs, ., by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO'))

# e.g.
bundy_ph <- dplyr::filter(horizon_layer, PROJECT_CODE == 'GCL' & 
                              FIELD_TEST_TYPE == 'PH' & FIELD_TEST_METHOD == 1 &
                              HORIZON_NO == 1 & FIELD_TEST_NO == 1)

ggplot() +
  geom_sf(data = bundy_ph, aes(colour = FIELD_TEST_VALUE), pch = 16, size = 2,
          show.legend = 'point') +
  scale_colour_brewer(palette = 'RdBu') +
  ggtitle("Surface pH", subtitle = "SALI Project GCL") +
  labs(colour = 'pH') +
  theme_minimal()
```

#### Vegetation Layer

Requires reconciling VCM, VST, VSP plus veg data from OBS

```{r veglayer}
veg_data <- sit_data[c(7, 17:19)]

veg_data <- lapply(veg_data, function(x) {
  dplyr::select(x, -matches('CREAT'), -matches('UPDAT')) 
  })

veg_data[['OBS']] <- veg_data[['OBS']] %>% 
  dplyr::select(PROJECT_CODE, SITE_ID, OBS_NO,
                FOREST_TYPE, matches('RAINFOR'), SCLEROPHYLL)

veg_data[['VCM']] <- veg_data[['VCM']] %>% 
  rename(VEG_COMM_NAME = NAME, VEG_COMM_DESCRIPTION = DESCRIPTION) %>% 
  # not in use 
  dplyr::select(-REF_VEG_COMM_NO)

vegetation_layer <- reduce(veg_data, full_join) %>% 
  right_join(locs, ., by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO'))

```

#### Notes

Just needs to be spatialised

```{r noteslayer}
notes_layer <- right_join(locs, sit_data[['NOT']],
                          by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) 
  

```

#### Samples

As for notes

```{r sampslayer}
samples_layer <- right_join(locs, sit_data[['SAM']],
                            by = c('PROJECT_CODE', 'SITE_ID', 'OBS_NO')) %>% 
  dplyr::select(-matches('CREAT'), -matches('UPDAT')) 

```

### Assemble

```{r flatsvc}
flat_service <- list(sites_layer, obs_layer, horizon_layer, vegetation_layer,
                     samples_layer, notes_layer)
names(flat_service) <- c('soil_sites', 'soil_observations', 'soil_horizons',
                         'vegetation', 'soil_samples', 'notes')
saveRDS(flat_service, 
        file.path(getwd(), 'DERIVED_DATA', 'SALI_Morph_Spatial_flat.rds'))
# flat_service <- readRDS(file.path(getwd(), 'DERIVED_DATA', 'SALI_Morph_Spatial_flat.rds'))
```

Quite a bit smaller than the simple conversion.

### Output

Writing to FGDB for use in Arc:

```{r create_gdb, eval= FALSE}
# NB currently experiencing problems using reticulate and arcgisbinding in the
# same session - one prevents the other finding the license >:-(
test_gdb <- file.path(getwd(), 'DERIVED_DATA', 'TEST_02_FLAT.gdb')

if(!(dir.exists(test_gdb))) {
  # launch access to arcpy
  library(reticulate)
  arcpy3_dir <- 'C:/Program Files/ArcGIS/Pro/bin/Python'
  system2(file.path(arcpy3_dir, 'Scripts', 'proenv.bat'))
  # make an empty file geodatabase
  arcpy <- import('arcpy')
  arcpy$CreateFileGDB_management(file.path(getwd(), 'DERIVED_DATA'), 
                                 'TEST_02_FLAT.gdb')
}

```

```{r write_gdb, eval = FALSE}
library(arcgisbinding) # arcgis interface
arc.check_product()

purrr::map2(flat_service, names(flat_service), function(lyr, nm) {
  arc.write(file.path(test_gdb, nm), lyr, overwrite = TRUE)
})

```


***
